Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Some more changes in description of NodIO architecture

In the work presented in this paper, JavaScript has been used throughout the full
stack, which has the advantage of using the self-same lines of code for
writing an evolutionary algorithm for the
desktop or command line and browser clients, both based in the EC
library {\sf NodEO} \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time 
when writing code and leverages 
expertise acquired
% (Paloma) Not sure if "acquired" should be written before "expertise"... sounds better to me, anyways
 in the library itself.

@@

In general, a distributed volunteer-based evolutionary computation
system based on the browser is simply a client-server system
whose client is, or it can be, embedded in the browser via
JavaScript. Since JavaScript is  the only language that is present
across all browsers, the choice was quite clear. Even so, there might
be some issues with the performance of the language itself. This
 is
why we have made a comparison between JavaScript and other languages
\cite{2015arXiv151101088M}, resulting that, in fact, JavaScript can be
faster than compiled languages such as Scala and, in any case, provides
a performance that is comparable to other languages usually employed  % (Paloma) Adverb
in evolutionary computation and numerical optimization such as Python. 
% (Paloma) Bit of a too long sentence here. Why not changing the ; for a . ?
Compiled languages such as
Java or C might be, indeed, faster, but the fact % (Paloma) Too many "fact" words. There was another one in the previous sentence.
 that a single-user
performance is higher is more than offset. This is because the number
of spontaneous volunteers we can gather at whim, including people
using their mobile phones or tablets, is several orders of magnitude
higher, which is the objective in this kind of systems.
% (Paloma) This sentence needs re-writing
% "Other compiled languages such as Java or C might be, in fact, faster, but the fact that single-user performance is higher (than what?) is over the offset. This is because the number of spontaneous volunteers we can gather at whim, including people using their mobile phones or tablets, is several orders of magnitude higher, which is the objective in this kind of systems." (?)
 JavaScript is
quite a popular language nowadays, providing compilers from a number
of languages, many of them in their family (like CoffeeScript), but
many other outside, so researchers can, in fact, write their fitness
function in Python, Erlang, Scala, or even Java, and them compile it to
JavaScript. For completness, an experiment comparing the performance of
a JavaScript implementation versus Java and Matlab is included in section
\ref{sec:w2}.

@@

Two kind of threats are possible in this environment; first, since the
API is open and it is an open source system, it is relatively easy to
sabotage the system, as indicated in \cite{domingues2007sabotage}, by
crafting a fake request which, for instance, assigns a fake fitness to
a particular chromosome
% (Paloma) Where is the second kind of threat?
. Another two different attacks are also
possible: a denial of service attack as well as penetration
testing. We have approached this threat in a pragmatic way: all
sources, client and server, for the application, are openly released. % (Paloma) Adverb
 We also release as open access the performed experiments \cite{DBLP:journals/corr/GuervosG15} and all data, once anonymized, as open data. This
builds trust between the user and the scientist, which indeed has an
impact on the performance of the system, since there is no need to
include cheating checks or other functions that would degrade the
performance of the system. This socio-technical approach is, indeed,
coherent with our whole approach to the study of a distributed volunteer
 system which can only be studied by approaching it from the
technical as well as the social, as in social media, angle.

@@

Thus, {\sf NodIO} architecture has two tiers:\begin{enumerate}
\item A REST server, that is, a server that includes several {\em
  routes} 
  that can be called for storing and retrieving information (the CRUD cycle:
  create, request, update, and delete) from the server. 
  A JSON data format is used for the communication between 
  clients and the server. There are two kinds of information:
  {\em problem} based, that is, related to the
  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
  GET}ing a random chromosome from it, and {\em information} related
  to the performance and state of the experiments. It also performs logging
  duties, but is basically a very lightweight and high performance
  data store \cite{jj:idc:lowcost}.
  The server has the capability to
  run a single experiment, storing the chromosomes in a data structure
  that is reset when the solution is found.
\item A client that includes the evolutionary algorithm as
  JavaScript code embedded in a web page that displays graphs and some
  additional links and information on the experiment. This code runs
  an evolutionary algorithm {\em island} starting with a random
  population and then sending, every 100 generations, the best individual
  back to the server (via a {\tt PUT} request) and requesting a random
  individual from the server (via a {\tt GET} request).
\end{enumerate}

>>>>

In the work presented in this paper, JavaScript has been used throughout the full
stack, which has the advantage of using the self-same lines of code for
writing an evolutionary algorithm for the
desktop or command line and browser clients, both based in the EC
library {\sf NodEO} \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time 
when writing code and leverages 
expertise acquired in the library itself.

@@

In general, a distributed volunteer-based evolutionary computation
system based on the browser is simply a client-server system
whose client is, or it can be, embedded in the browser via
JavaScript. Since JavaScript is  the only language that is present
across all browsers, the choice was quite clear. Even so, there might
be some issues with the performance of the language itself. This
 is
why we have made a comparison between JavaScript and other languages
\cite{2015arXiv151101088M}, resulting that, in fact, JavaScript can be
faster than compiled languages such as Scala and, in any case, provides
a performance that is comparable to other languages usually employed  
in evolutionary computation and numerical optimization such as Python. 
% (Paloma) Bit of a too long sentence here. Why not changing the ; for a . ?
Compiled languages such as
Java or C might be, indeed, faster, but the fact % (Paloma) Too many "fact" words. There was another one in the previous sentence.
 that a single-user
performance is higher is more than offset. This is because the number
of spontaneous volunteers we can gather at whim, including people
using their mobile phones or tablets, is several orders of magnitude
higher, which is the objective in this kind of systems.
% (Paloma) This sentence needs re-writing
% "Other compiled languages such as Java or C might be, in fact, faster, but the fact that single-user performance is higher (than what?) is over the offset. This is because the number of spontaneous volunteers we can gather at whim, including people using their mobile phones or tablets, is several orders of magnitude higher, which is the objective in this kind of systems." (?)
 JavaScript is
quite a popular language nowadays, providing compilers from a number
of languages, many of them in their family (like CoffeeScript), but
many other outside, so researchers can, in fact, write their fitness
function in Python, Erlang, Scala, or even Java, and them compile it to
JavaScript. For completness, an experiment comparing the performance of
a JavaScript implementation versus Java and Matlab is included in section
\ref{sec:w2}.

@@

Two kind of threats are possible in this environment; first, since the
API is open and it is an open source system, it is relatively easy to
sabotage the system, as indicated in \cite{domingues2007sabotage}, by
crafting a fake request which, for instance, assigns a fake fitness to
a particular chromosome
% (Paloma) Where is the second kind of threat?
. Another two different attacks are also
possible: a denial of service attack as well as penetration
testing. We have approached this threat in a pragmatic way: all
sources, client and server, for the application, are openly released. 
 We also release as open access the performed experiments \cite{DBLP:journals/corr/GuervosG15} and all data, once anonymized, as open data. This
builds trust between the user and the scientist, which indeed has an
impact on the performance of the system, since there is no need to
include cheating checks or other functions that would degrade the
performance of the system. This socio-technical approach is, indeed,
coherent with our whole approach to the study of a distributed volunteer
 system which can only be studied by approaching it from the
technical as well as the social, as in social media, angle.

@@

Thus, {\sf NodIO} architecture has two tiers:\begin{enumerate}
\item A REST server, that is, a server that includes several {\em
  routes} 
  that can be called for storing and retrieving information (the CRUD cycle:
  create, request, update, and delete) from the server. 
  A JSON data format is used for the communication between 
  clients and the server. There are two kinds of information:
  {\em problem} based, that is, related to the
  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
  GET}ing a random chromosome from it, and {\em information} related
  to the performance and state of the experiments. It also performs logging
  duties, but they are basically a very lightweight and high performance
  data storage \cite{jj:idc:lowcost}.
  The server has the capability to
  run a single experiment, storing the chromosomes in a data structure
  that is reset when the solution is found.
\item A client that includes the evolutionary algorithm as
  JavaScript code embedded in a web page that displays graphs, some
  additional links, and information on the experiment. This code runs
  an evolutionary algorithm {\em island} starting with a random
  population, then it sends, every 100 generations, the best individual
  back to the server (via a {\tt PUT} request), and requests a random
  individual from the server (via a {\tt GET} request). % (Paloma) Re-writing because there are 3 steps (island creation -> PUT -> GET)
\end{enumerate}
